First and foremost I want to sincerely thank Tom, my PhD advisor, who gave me the
opportunity to pursue the doctoral program with his guidance. Tom is a fantastic
advisor, a statement that I have been saying time and again especially new
students seeking an advisor. In Indian culture we have a special place and
respect for the ``guru'' and I feel that I was fortunate enough to find Tom and
to be able to associate with him as my ``guru''. Tom is a highly energetic,
optimistic, professional, and extremely knowledgeable person. His advising style
is very adaptive and he let's each student learn and work at his/her own pace.
He does not create work pressure, but if you are going off tracks he will for
sure warn you. He is an extremely adaptive and flexible advisor be it technically,
administratively, or logistically. Because of his support I was able to explore
different positions in industry and understand the contrasting but complementary
research approaches in academia versus industry. With his guidance I have learnt
to deal with rejections and failures with a positive attitude, which is an
instrumental medicine for PhD journey. No matter how many rejections you get,
his philosophy of keep trying to improve and doing excellent quality research is
what I found most useful as his student. I am thankful to University of
Michigan, CSE department to give me a wonderful opportunity of associating with
Tom, as my advisor.

I also want to take this chance to thank all of my collaborators for their
contributions in making my ideas more concrete and provide me the perspective
that I lacked before. I especially want to mention about the guidance and
support I got from David Nellans. Dave was my internship mentor at NVIDIA.  His
approach of conducting product influential research helped me understand impact
of research on products and business utility for a company.  Dave is also a very
good friend who boosts my moral and motivates me to keep doing good work.
Andr\'es Lagar-Cavilla is another important name in my technical career,
because of whom I got interested in Linux and operating systems. His commitment
to help and support his peers inspires me. His help during my internship at
Google was the stepping stone for my success in my internship project.

I think coming into the PhD program was one of the best decisions of my life. I
got many opportunities to travel in the last six years, exploring different
countries, cultures, and lifestyles. Travel gave me a perspective to meaning of
life. I could go out of my comfort zone, experience people's generosity,
overcome difficulty of not knowing whereabouts of a place. I realized that world
is a much bigger and beautiful place and that experience matters more than
success. Travel inspired me to explore new avenues in my research, become
comfortable at switching projects and keep learning new things.

My parents -- Madhu Agarwal and Bal Krishna Agarwal -- are the reason whatever I
am today. Their constant guidance and support in life has been instrumental in
my writing this acknowledgement section today. They have provided me the shade
and comfort at each and every single step in my life even when I didn't ask for
it. They always understand what I need and are always there for me. The feeling
of having them no matter what happened in my professional or personal life is
the biggest gift I have gotten in my life. When things were not going the way I
wanted them to go for very long time and I was doubting myself all the time,
mummy papa were the one who always filled me with confidence and encouraged me
to try even harder. I have called them several times late at nights to talk about
how worried I am, how things are not happening even though I am doing all I can,
they are always there to listen to me quietly and help me move forward. I can't
imagine doing anything without their support. Thanks is a small word for what my
parents have done for me. I can just say that I am lucky enough to have them in
my life. Mummy--Papa this dissertation is your hard work and you confidence in
me to be able to sincerely work and contribute to the world.

Next, I want to talk about my brother, Shyam Mohan Agarwal. He is the person who
loves me so much that at times I think what did I do to receive that much love.
He is the one because of whom I am able to fulfill my dream of studying in
the best institutes of knowledge. His encouragement and belief in me is why I
have the courage to try out new things fearlessly. He looked something in me
since childhood and he knew that I would be a scholar. It has been like he has
foreseen what I will become and then gave me a push to go and get it. With every
achievement of mine it feels like he actually won something. He is the one who
enjoys my success the most. Without him experiencing the joy of what I do my
work is incomplete. He is a true inspiration to me, without whom I am not sure
how I would have a direction to pursue something that I love most doing. I also
want to thank my sister-in-law, Khyati Agarwal, who has been so nice to
accommodate all of us in her life and continued to make our family loving and
happy.

I have also been very fortunate to have come across several fantastic people and be
friends with them. Gaurav, Pooja, Kataria, Gauri, Tanvi, Mohit, Kunal, Pulkit
are some names very close to my heart that I know will come for me if I need
them at any point in life. They have challenged me in several ways, making me a
better person with every interaction. Countless nights we have just sat together
and just discussed some weird mystery of life or just cribbed about how things
are unfair. These people are treasures I have found in my life and have made my
life so comfortable and loving. They have played a significant role by keeping
me sane during the most self-doubting periods of PhD journey. All of my seniors
Gaurav Chadha, Ankit Sethia, Daya Khudia, and Abhayendra Singh are some of the
people who helped me acclimatize in Ann Arbor and the department during my
foundational years of my PhD.

I want to thank my labmates for being very supportive all throughout my graduate
school. Steve Pelley has been a mentor to me and I always go to him to ask for
all sorts of career advice. Faissal Sleiman is a hard working person who always
inspires me to do the right thing. Akshitha, Vaibhav, Amlan, Aasheesh, Jeff all
have been some of the fantastic, smart, and hard working individuals I have come
across, who are like extended family to me. I also want to acknowledge CSE staff
members Dawn Freysinger, Lauri Johnson, Karen Liska, Denise Duprie, and Stephen
Reger who have helped in every possible way whenever I needed
administrative/logistics support from the department.

Finally, I would like to mention the most significant person Subhasis Das,
because of whom I have been able to survive the graduate school. He is the most
amazing person that has happened to me in this life after having blessed with a
fantastic family. This is the person without whom last five years would have been
extremely hard. Even after living miles away he can bring smile on my face in
seconds and make everything look good again. His presence in my life is
priceless and I can't thank God enough for bringing him in my life.  I will
cherish all the fun we had discussing technical stuff, watching Netflix
together and a getaway from cold Ann Arbor. I have learnt a lot from Subhasis's
technical approach and his philosophy that at times you got to try million
things before something actually starts working or is meaningful. Thanks so much
for being part of my life, without you this journey will be incomplete.

